# 4 Analysis

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```
__TO DO's__

* Answers to the research questions
* Different methods considered
* Competing approaches
* Justifications

## 4.1 RQ1

## 4.2 RQ2

## 4.3 RQ3

In comparing regions in Switzerland, which areas show higher or lower adoption of electric vehicles, and how does this regional adoption align or vary with external factors like oil price changes, political opinions, and demographic shifts?

```{r}
head(df_v_map)
```


```{r}
# Assume you have two dataframes: ev_adoption_data and demographic_data
# ev_adoption_data has columns 'Year' and 'EV_Count'
# demographic_data has columns 'Year', 'GenerationZ', 'Millennials', 'GenerationX', 'BabyBoomers'

# Summing up the total count of electric vehicles by year
total_ev_count <- df_ele %>%
                  group_by(Year) %>%
                  summarise(Total_EV_Count = sum(EV_Count))

# Calculating total population for each year in demographic data
demographic_data <- demographic_data %>%
                    mutate(Total_Population = GenerationZ + Millennials + GenerationX + BabyBoomers)

# Merging EV adoption data with demographic data
combined_data <- merge(total_ev_count, demographic_data, by = "Year")

# Calculating EV adoption rate per capita
combined_data$EV_Adoption_Rate_Per_Capita <- combined_data$Total_EV_Count / combined_data$Total_Population

# View the combined data
head(combined_data)

```


```{r}
# Assuming 'combined_data' is your dataframe with the following columns: 'Year', 'Total Population', 'Total EV Count'
combined_data$EV_Adoption_Rate_Per_Capita <- combined_data$Total_EV_Count / combined_data$Total_Population

ggplot(combined_data, aes(x = Year, y = EV_Adoption_Rate_Per_Capita)) +
  geom_line() +
  theme_minimal() +
  ggtitle("EV Adoption Rate Per Capita in Switzerland (2005-2009)") +
  xlab("Year") +
  ylab("EV Adoption Rate Per Capita")

```


```{r}
library(reshape2)

# Assuming 'combined_data' has columns for generational proportions and EV adoption rate
# Let's say the relevant columns are 'GenZ_Proportion', 'Millennials_Proportion', 'GenX_Proportion', 'Boomers_Proportion', 'EV_Adoption_Rate_Per_Capita'

correlation_data <- combined_data[, c('GenZ_Proportion', 'Millennials_Proportion', 'GenX_Proportion', 'Boomers_Proportion', 'EV_Adoption_Rate_Per_Capita')]
correlation_matrix <- cor(correlation_data)

melted_correlation_matrix <- melt(correlation_matrix)
ggplot(data = melted_correlation_matrix, aes(Var1, Var2, fill = value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() +
  ggtitle("Correlation Matrix between Generational Proportions and EV Adoption Rate Per Capita") +
  xlab("") +
  ylab("")

```

## 4.4 RQç

## 4.5  To what extent does the evolution in the availability of charging stations exert an influence on the adoption of electric     vehicles in Switzerland?
```{r}
# First, let's merge the df_v and df_charge_number_CH data sets, and we will look at Fuel: Electricity and Plug-in Hybrid

df_v_electric_total_ch <- df_v %>%
  filter(Fuel == c("Electricity", "Plug-in hybrid"), VehicleType == "Passenger car", Location == c("Switzerland","Confederation")) %>%
  select(Date, Count)

sum_by_year <- df_v_electric_total_ch %>%
  group_by(Year = lubridate::year(Date)) %>%
  summarise(Total_Count = sum(Count))


# Convert year to a common format for merging
sum_by_year <- sum_by_year %>%
  mutate(year = as.Date(paste0(Year, "-01-01")))

# Merge the datasets based on the "year" column
merged_v_charge <- left_join(sum_by_year, df_charge_number_CH, by = c("year" = "year"))

# cleaning merged data set
merged_v_charge <- merged_v_charge %>%
  filter(Year > "2011") %>%
  select(Year, Total_Count, powertrain, value)

names(merged_v_charge)[names(merged_v_charge) == "Total_count"] <- "EVs"
colnames(merged_v_charge)[colnames(merged_v_charge) == "value"] <- "Charging station"

# Summing Powertrain together
merged_v_charge <- merged_v_charge %>%
  group_by(Year, Total_Count) %>%
  summarise(Count = sum(`Charging station`))

# Checking the correlation
corr_charge_ev <- cor(merged_v_charge$Total_Count, merged_v_charge$Count)

# their correlation is 0.957, almost perfectly correlated (no suprise here)

# Checking for lagged correlation

lags_to_explore <- 1:3

lagged_correlation <- function(data, lag) {
  data %>%
    mutate(Count_Lagged = lag(Count, n = lag, default = NA)) %>%
    summarise(Correlation = cor(Total_Count, Count_Lagged, use = "complete.obs"))
}

# Calculate lagged correlations for each lag
lagged_correlations_df <- data.frame(Lag = lags_to_explore) %>%
  rowwise() %>%
  mutate(Correlation = lagged_correlation(merged_v_charge[, -1], Lag)$Correlation)

# Print the results
print("Original Correlation:")
print(corr_charge_ev)

print("Lagged Correlations:")
print(lagged_correlations_df)


# Now we formulate the following Hypothesis

# H0: new charging station increase EV adoption vs. H1: new charging station does not increase EV adoption

# Check these hypotheses with a simple linear regression
linear <- lm(Total_Count ~ Count, data = merged_v_charge)

# Poisson Test
poisson_model <- glm(Total_Count ~ Count, family = poisson, data = merged_v_charge)

# Set up the layout using mfrow
par(mfrow = c(1, 2))  # 1 row, 2 columns

# Plotting for Simple Linear Regression
plot(linear, 1, main = "LM Residuals vs Fitted")
plot(linear, 2, main = "LM Normal Q-Q Plot")
plot(linear, 3, main = "LM Scale-Location Plot")
plot(linear, 5, main = "LM Residuals vs Leverage")

# Plotting for Poisson Regression
plot(poisson_model, which = 1, main = "Poisson Residuals vs Fitted")
plot(poisson_model, which = 2, main = "Poisson Normal Q-Q Plot")
plot(poisson_model, which = 3, main = "Poisson Scale-Location Plot")
plot(poisson_model, which = 5, main = "Poisson Residuals vs Leverage")

# Reset the layout
par(mfrow = c(1, 1))
```
The year-on-year correlation is the highest, the lagged correlation diminishes. We could suggest hat the availability of charging stations and the new registration of EVs go hand-in-hand, and that the availability of new charging station does not create a demand of new EVs by itself. Correlation does not imply causation, while we see a relationship, we can't conclude that charging stations directly cause changes in electric vehicle adoption only with a Correlation analysis

With both a linear regression and a Poisson-test. We find evidence of statistically significant relationship between the count of available charging station and the count of electric vehicles registered. We have Prediction variable/coefficient of 2.69 x 10^-4 and 3.34 respectively. And a p-val < 0.005 for both. However, it is important to remind ourselves that these variables have a bidirectional / mutual influence, beyond the scope of what our analysis shows. The relationship is not strictly unidirectional and therefore, it is hard to conclude anything without further domain-knowledge and context-specific information

