# 4 Analysis

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```
__TO DO's__

* Answers to the research questions
* Different methods considered
* Competing approaches
* Justifications

## 4.1 RQ1

## 4.2 RQ2

## 4.3 RQ3

In comparing regions in Switzerland, which areas show higher or lower adoption of electric vehicles, and how does this regional adoption align or vary with external factors like oil price changes, political opinions, and demographic shifts?

```{r}
head(df_v_map)
```


```{r}
# Assume you have two dataframes: ev_adoption_data and demographic_data
# ev_adoption_data has columns 'Year' and 'EV_Count'
# demographic_data has columns 'Year', 'GenerationZ', 'Millennials', 'GenerationX', 'BabyBoomers'

# Summing up the total count of electric vehicles by year
total_ev_count <- df_ele %>%
                  group_by(Year) %>%
                  summarise(Total_EV_Count = sum(EV_Count))

# Calculating total population for each year in demographic data
demographic_data <- demographic_data %>%
                    mutate(Total_Population = GenerationZ + Millennials + GenerationX + BabyBoomers)

# Merging EV adoption data with demographic data
combined_data <- merge(total_ev_count, demographic_data, by = "Year")

# Calculating EV adoption rate per capita
combined_data$EV_Adoption_Rate_Per_Capita <- combined_data$Total_EV_Count / combined_data$Total_Population

# View the combined data
head(combined_data)

```


```{r}
# Assuming 'combined_data' is your dataframe with the following columns: 'Year', 'Total Population', 'Total EV Count'
combined_data$EV_Adoption_Rate_Per_Capita <- combined_data$Total_EV_Count / combined_data$Total_Population

ggplot(combined_data, aes(x = Year, y = EV_Adoption_Rate_Per_Capita)) +
  geom_line() +
  theme_minimal() +
  ggtitle("EV Adoption Rate Per Capita in Switzerland (2005-2009)") +
  xlab("Year") +
  ylab("EV Adoption Rate Per Capita")

```


```{r}
library(reshape2)

# Assuming 'combined_data' has columns for generational proportions and EV adoption rate
# Let's say the relevant columns are 'GenZ_Proportion', 'Millennials_Proportion', 'GenX_Proportion', 'Boomers_Proportion', 'EV_Adoption_Rate_Per_Capita'

correlation_data <- combined_data[, c('GenZ_Proportion', 'Millennials_Proportion', 'GenX_Proportion', 'Boomers_Proportion', 'EV_Adoption_Rate_Per_Capita')]
correlation_matrix <- cor(correlation_data)

melted_correlation_matrix <- melt(correlation_matrix)
ggplot(data = melted_correlation_matrix, aes(Var1, Var2, fill = value)) + 
  geom_tile() +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Pearson\nCorrelation") +
  theme_minimal() +
  ggtitle("Correlation Matrix between Generational Proportions and EV Adoption Rate Per Capita") +
  xlab("") +
  ylab("")

```

## 4.4 RQç

## 4.5 RQ5

